---
title: "BellaBeat_casestudy"
author: "Minty Vo"
date: "29/01/2022"
output: html_document
---
*Welcome to my Bellabeat data analysis case study*

## About the company Bellabeat 
Bellabeat, is a high-tech manufacturer of health-focused products for women. Bellabeat is a successful small company, and they have the potential to become a larger player in the global smart device market. Urška Sršen and Sando Mur founded Bellabeat, a high-tech company that manufactures health-focused smart products. Collecting data on activity, sleep, stress, and reproductive health has allowed Bellabeat to empower women with knowledge about their own health and habits. Since it was founded in 2013, Bellabeat has grown rapidly and quickly positioned itself as a tech-driven wellness company for women. 

## Scenario of the study 
In this study, I will focus on one of Bellabeat’s products and analyze smart device data to gain insight into how consumers are using their smart devices. The insights will then help guide marketing strategy for the company

 I will follow the steps of the data analysis process: ask, prepare, process, analyze, share, and act. 

## Ask
 **Business task:** 
 
Smart devices are a big part of people’s everyday life. As a smart device manufacturer, Bellabeat can benefit from learning the trend of smart device usage and make data-driven business strategies to explore opportunities for growth.

 **Questions for the analysis:**
 
  1. What are some trends in smart device usage?
  
  2. How could these trends apply to Bellabeat customers?
  
  3. How could these trends help influence Bellabeat’s marketing strategy?
  
## Prepare

* Where is the data stored?

Data is publicly available on Kaggle: FitBit Fitness Tracker Data. 

* How is the data organized?

Data is stored in 18 csv files.

* Are there issues with bias or credibility in this data? Does the data ROCCC?

This Kaggle data set contains personal fitness tracker from 30 fitbit users. They were generated by respondents to a distributed survey via Amazon Mechanical Turk between 12.04.2016-12.05.2016.

 Reliable — LOW — Not reliable as it only has 30 respondents 
 Original — LOW — Third party provider (Amazon Mechanical Turk) 
 Comprehensive — MED — Parameters match most of Bellabeat product´s parameters  Current — LOW — Data is 6 years old and may not be relevant 
 Cited — LOW — Data collected from third party, hence unknown 
 “Overall, this dataset is considered “bad quality data” and it is not recommended to produce business recommendations based on this data”

* How are you addressing licensing, privacy, security, and accessibility?

30 eligible Fitbit users consented to the submission of personal tracker data, including minute-level output for physical activity, heart rate, and sleep monitoring. 

* How did you verify the data’s integrity?

As data is collected in a survey, we are unable to ascertain its integrity or accuracy.

* How does it help you answer your question?

This data explores smart device users’ daily habits. It includes information about daily activity, steps, sleep habits and heart rate, that can be used to explore users’ habits and find some trends.

* Are there any problems with the data?

Data was collected 6 years ago in 2016. Users’ daily activity, fitness and sleeping habits, diet and food consumption may have changed since then. Data may not be timely or relevant. Sample size of 30 FitBit users is not representative of the entire fitness population.

## Process
**Downloading the data**
 
The dataset was generated by respondents to a distributed survey via Amazon Mechanical Turk between 03.12.2016 and 05.12.2016. And include 18 CSV files. 

**Loading packages** 
I am going to install and load some R packages that will help in my analysis 

```{r}
# install.packages("tidyverse")
# install.packages("lubridate")
# install.packages("dplyr")
# install.packages("ggplot2")
# install.packages("tidyr")
# install.packages("here")
# install.packages("skimr")
# install.packages("janitor")
# install.packages("sqldf")
library(tidyverse)
library(lubridate)
library(dplyr)
library(ggplot2)
library(tidyr)
library(here)
library(skimr)
library(janitor)
library(sqldf)
```

Now I import all the dataset on RStudio and  use head(), colnames(), str() to quickly review the data 

```{r}
setwd("/Users/minty/Developer/Data_analysis_project")
daily_activity <- read_csv("dailyActivity_merged.csv")

head(daily_activity)

colnames(daily_activity)

str(daily_activity)


daily_calories <- read_csv("dailyCalories_merged.csv")

head(daily_calories)

colnames(daily_calories)


daily_intensities <- read_csv("dailyIntensities_merged.csv")

head(daily_intensities)

colnames(daily_intensities)


hourly_intensities <- read_csv("hourlyIntensities_merged.csv")

head(hourly_intensities)

colnames(hourly_intensities)


daily_steps <- read_csv("dailySteps_merged.csv")

head(daily_steps)

colnames(daily_steps)


heart_rate_sec <- read_csv("heartrate_seconds_merged.csv")

head(heart_rate_sec)

colnames(heart_rate_sec)


minute_METs <- read_csv("minuteMETsNarrow_merged.csv")

head(minute_METs)

colnames(minute_METs)


sleep_day <- read.csv("sleepDay_merged.csv")

head(sleep_day)

colnames(sleep_day)


weight_log <- read_csv("weightLogInfo_merged.csv")

head(weight_log)

colnames(weight_log)
```
 
 I’m going to Process, Clean and Organize the dataset for analysis. 
 I spotted some problems with the timestamp data. So before analysis, I need to convert it to date time format and split to date and time. 

```{r }
daily_activity$ActivityDate=as.POSIXct(daily_activity$ActivityDate, format="%m/%d/%Y", tz=Sys.timezone())
daily_activity$date <- format(daily_activity$ActivityDate, format = "%m/%d/%y")

sleep_day$SleepDay=as.POSIXct(sleep_day$SleepDay, format="%m/%d/%Y %I:%M:%S %p", tz=Sys.timezone())
sleep_day$date <- format(sleep_day$SleepDay, format = "%m/%d/%y")
sleep_day$month <- format(sleep_day$SleepDay, format = "%B")

heart_rate_sec$Time=as.POSIXct(heart_rate_sec$Time, format="%m/%d/%Y %I:%M:%S %p", tz=Sys.timezone())
heart_rate_sec$date <- format(heart_rate_sec$Time, format = "%m/%d/%y")
heart_rate_sec$time <- format(heart_rate_sec$Time, format = "%H:M%:S%")

weight_log$Date=as.POSIXct(weight_log$Date, format="%m/%d/%Y %I:%M:%S %p", tz=Sys.timezone())
weight_log$date <- format(weight_log$Date, format = "%m/%d/%y")
weight_log$time <- format(weight_log$Date, format = "%H:%M:%S")

hourly_intensities$ActivityHour=as.POSIXct(hourly_intensities$ActivityHour, format="%m/%d/%Y %I:%M:%S %p", tz=Sys.timezone())
hourly_intensities$time <- format(hourly_intensities$ActivityHour, format = "%H:%M:%S")
hourly_intensities$date <- format(hourly_intensities$ActivityHour, format = "%m/%d/%y")

```

```{r }

```